---
title: "AssayR"
author: "jimi.wills@ed.ac.uk"
date: "10 June 2017"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{AssayR}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}

---

<!--
don't forget to build this vignette with 
devtools::build_vignettes("AssayR") 
-->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AssayR

AssayR provides a small number of functions for simple and
rapid assay of high resolution LCMS data, especially isotopically
labelled tracer experiments.

Given a table of desired compounds, in a specific format, AssayR will extract
chromatograms for those compounds, and interactively allow you to optimise peak
detection.  Peak detection is performed on sthe maximal trace for the whole data
set, including any isotopes you've defined, and so the approach is immune to 
missing peaks.

The output is a table of results (peak areas).  With a labelling experiment, 
there's also graphical output in the form of stacked bars to summarise the 
information (both percentage and aboslute).

To install AssayR (first install R and optionally RStudio) run the following
commands in R:

    install.packages(c("RColorBrewer", "reshape2", "stringr","mzR"))
    install.packages("https://gitlab.com/jimiwills/assay.R/raw/master/AssayR_0.1.2.tar.gz", repos = NULL, type = "source")

If you have 
not yet installed R, please visit [CRAN](https://cran.r-project.org/).  And I 
heartily recommend 
[RStudio Desktop](https://www.rstudio.com/products/RStudio/#Desktop).

## Summary of use


```{r usage, eval=FALSE}
# change directory to where your raw files are
setwd(path.to.raw.files)
# you can also do this in the RGui or RStudio menu.

# convert raw files to mzML if they are not already
msconvert_all()
# this extract pos and neg and puts them in mzMLpos and mzMLneg
# directories

# generate TICs/EICs/XICs from config file
path.to.tics <- run.config.tics(path.to.config, path.to.mzML)
# more about the config file below

# analyse the peaks (interactively)
results <- run.config.peaks(path.to.config, path.to.tics)
# this is interactive, more below.
# you could write the result to csv
# write.csv(results, "results.csv")

# rearrange and output the results, and output plots of the
# isotopes:
assay.plotter(results)
# check the current working directory for outputs

```

## Config file

There are example config files with this package.  To access them use a command
like this:

    system.file("extdata/configs/pyrimidine.tsv", package = "AssayR")

And to like the files available:

    list.files(system.file("extdata/configs", package = "AssayR"))


The format should be a tsv, as readable by `read.delim`, and the file
should have the following columns:

* target
    name of the target channel (gets used for naming data)
* name
    name of the compound (gets used for labelling graphs)
* mz
    monoisotopic m/z value for this compound
* ppm
    m/z tolerance in ppm for this compound
* rt.min
    minimum retention time to be considered (can be modified interactively)
* rt.max
    maximum retention time to be considered (can be modified interactively)
* seconds
    width of pick-detection wave in seconds (can be modified interactively)
* threshold
    minimum required intensity (can be modified interactively)
* known alternatives
    reserved for future use
* C13
    in labelled experiment, maximum possible number of 13C in compound
* N15
    in labelled experiment, maximum possible number of 15N in compound
* H2
    in labelled experiment, maximum possible number of 2H in compound
* event
    in mass spec method with multiple events (scans), indicates in which event
    to find the m/z of this compound
* interactive
    values: "yes" or "done", indicates whether interactive mode is 
    required for this definition
* comment
    free text.  I use this for sorting my compounds in pathway order.
    
    
## more information?

If you think something's missing, please get in contact and I'll update this
document so everybody can benefit.  Thanks!
